Comparative study of Dask and Spark for neuroimaging pipelines

### 2019-06-11 latex/acl2019.tex
Small fix

\section{Introduction}
% Context
With the increasing number of data available in neuroimaging \citep{ALFAROALMAGRO:18,
UKBioBank:18} the processing of Big Data is becoming critical. Frameworks like Nipype
\citep{Nipype:11} are usually used to create neuroimaging pipelines however it is
worth considering the use of general purpose Big Data frameworks
\citep{Hayot-Sasson:17}. In that research, Spark \citep{Spark:16} was a natural
choice as it is the state-of-the-art framework for Big Data. However, as our
laboratory works mostly with Python, we want to know if Dask \citep{Dask:15}, a
Python written framework, could bring better or similar performances while
facilitating usage.

@@

% Spark
On the one hand, we have Spark which provides a high-level API. This allows the
scheduler to perform more optimizations which makes it well suited for neuroimaging
applications that often require lengthy executions. However, Spark's code base is in
Scala which theoretically can lead to slow down in execution due to a required
serialization. Moreover, while Spark's API is flexible and allows most
implementations, it differs from the ones seen in the Python's ecosystem.

@@

% Methods used (summary)
The project aims to compare the state-of-the-art general purpose framework Spark with
the newcomer Dask. We decide to compare the performance of Spark and Dask on custom
incrementation pipeline to simplify the effect of the algorithm on the comparison.
Then we access the frameworks on two real-life applications: (1) histogram of the
voxels intensity in a 3d images (2) BIDS example.

% Implications of our research
The result from our project help in assessing if Dask is a good choice to build
neuroimaging pipelines.

%%%%% MATERIAL AND METHODS %%%%%%
\section{Material and Methods}
%% Material
We compare Dask(\textbf{v1.1.4}) and Spark(\textbf{v2.4.0}) based on two metrics
their scheduling and their makespan performance. All scripts used to perform our
experiments are available at
\href{https://github.com/mathdugre/SOEN-499-Project}{https://github.com/mathdugre/SOEN-499-Project}.
For our experiments we used Compute Canada Cloud where each instance has 8 cores and
30GB RAM. The system was dedicated.

%% Incrementation
\subsection{Incrementation}
As our first experiment to benchmark the frameworks we used a simple incrementation
pipeline (see Algorithm \ref{alg:incrementation}). It loads chunks from the 75GB
BigBrain\citep{ds052:01} image, increment its value by 1 and save the result to an
nfs as a Nifti image. The incrementation allows to reduce caching effect of the data.
This experiment allows us to study the behaviour of the frameworks when all input are
processed independently; i.e. each task in the graph only depend on the previous one
(see \ref{fig:tg-inc}).

@@

% Histogram
\subsection{Histogram}
For our second experiment we find the frequency of the voxels intensity in BigBrain
and save them as a tuple in a file. In this experiment the end results depends on all
of the input results (see \ref{fig:tg-histo}). This allow us to study the behaviour
of the framework when task are dependent. Also this application require data
shuffling thus inter worker communication. In this application we decided to omit the
dask futures api as it is less realistic for this use case.

@@

% BIDS Example
\subsection{BIDS Example}
Our last experiment compare both frameworks on the BIDS Example algorithm.
% Specify more details and the data set used

@@

%% Incrementation
For the incrementation experiment, we
%% Node
first vary the number of instances in our cluster environment. We test with 1, 2, 4,
8, and 16 instances. From this we can evaluate the level of parallelism offered by
both framework.

@@

%% Splits
Finally, we split the BigBrain in different number of chunks. We split it in 30, 125
and 750 chunks. From that we can evaluate the framework with task of different size;
around 2.5GB, 0.6GB and 0.1GB, respectively.

%% Baseline
Our baseline for all experiment is 10 iterations, 4 sleep delay, 125 chunks running
on 8 instances.

@@

%%% INCREMENTATION %%%
%% Instances
\subsection{Experiment 1: Number of instances}
Figure \ref{fig:inc_node}(a) shows that the incease in performance is not
proportional to the number of worker which differ form what we would expect; it even
decrease when we reach 8 instances (64 workers). This is explained by an increase in
IO time corrolated to the number of workers (Figure \ref{fig:inc_node}(b)), as more
worker access the nfs simultaneously it increase the stress put on the system which
result in bottleneck, errors and slow down.

On the one hand, Dask seems better when
using a single worker. On the other hand, Spark makespan decrease faster for low
level of parallelism. However Dask quickly catch up at high level of parallelism.
This makes us think that the scheduler in Dask might be better

@@

%% Iterations
\subsection{Experiment 1: Number of iterations}
In Figure \ref{fig:inc_itr} we see that spark and the dask bag api are equivalent in
term of performance. The dask delayed api seems to perform well with low amount of
task but grow faster the spark and dask bag. We think this is because dask delayed
often schedule a block on different worker of the same instance which cause small
delay everytime due to inter worker communication. This is good when there is fewer
tasks as is make the IO between block go out of sync hence lowering the IO bottleneck
however as the number of tasks increases those small delays become significant. The
dask futures api seems to outperform all the other apis but this is most likely
because the tasks are not inter dependent thus this application benifits from the
less optimal but faster scheduling brought by dask futures.

@@

%% Sleep time
\subsection{Experiment 1: Sleep delay}
As Figure \ref{fig:inc_sleep}

@@

% Chunks
\subsection{Experiment 1: Number of splits}
In Figure \ref{fig:inc_splits} Spark is not compared for 30 chunks. This is because
spark has a 2GB limitation in the task size it can compute. Note from Figure
\ref{fig:inc_splits}(b) that the compute time increase when there is more splits.
This is expected because the sleep delay is constant throughout this experiment and
there are more compute task when the number of splits increases. We also note that in
Figure \ref{fig:inc_splits}(b) for 30 splits the dask bag api has a much lower
overhead time. This is because calculate the idle time of active worker only and the
dask bag api was only using one worker per split in comparison to the dask delayed
and dask futures api which were offloading the calculations on multiple workers on a
same instance. From Figure \ref{fig:inc_splits} we observe that when the dataset is
seperated in more (smaller) splits it saves IO time however it increases the overhead
by a similar amount; the inverse happens when we reduce teh number of splits.

>>>>

\section{Introduction}
% Context
With the increasing number of data available in neuroimaging \citep{ALFAROALMAGRO:18,
UKBioBank:18} the processing of Big Data becomes critical. Frameworks like Nipype
\citep{Nipype:11} are usually used to create neuroimaging pipelines however it is
worth considering the use of general purpose Big Data frameworks
\citep{Hayot-Sasson:17}. In that research, Spark \citep{Spark:16} was a natural
choice as it is the state-of-the-art framework for Big Data. However, as our
laboratory works mostly with Python, we want to know if Dask \citep{Dask:15}, a
Python written framework, could bring better or similar performances while
facilitating usage.

@@

% Spark
On the one hand, we have Spark which provides a high-level API. This allows the
scheduler to perform more optimizations which makes it well suited for neuroimaging
analysis that often require the usage of pipeline with multiple steps. However,
Spark's code base is in Scala which theoretically can lead to slow down in execution
due to a required serialization. Moreover, while Spark's API is flexible and allows
most implementations, it differs from the ones seen in the Python's ecosystem.

@@

% Methods used (summary)
The project aims to compare the state-of-the-art general purpose framework Spark with
the newcomer Dask. We decide to compare the performance of Spark and Dask on a custom
incrementation pipeline to simplify the effect of the algorithm on the comparison.
Then we assess the frameworks on two real-life applications: (1) histogram of the
voxels intensity in a 3d images (2) BIDS example.

% Implications of our research
The result from our project help in deciding if Dask is a good choice to build
neuroimaging pipelines.

%%%%% MATERIAL AND METHODS %%%%%%
\section{Material and Methods}
%% Material
We compare Dask(\textbf{v1.1.4}) and Spark(\textbf{v2.4.0}) based on two metrics
their scheduling and their makespan performance. All scripts used to perform our
experiments are available at
\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines}.
For our experiments we used Compute Canada Cloud where each instance has 8 cores,
30GB RAM, 10GB bandwidht and is dedicated.

%% Incrementation
\subsection{Incrementation} As our first experiment to benchmark the frameworks we
used a simple incrementation pipeline (see Algorithm \ref{alg:incrementation}). It
loads chunks from the 75GB BigBrain\citep{ds052:01} image, increment its value by 1 a
determined number of time and save the result to an nfs as a Nifti image.
Incrementting the data reduce the caching effect. This experiment allows us to study
the behaviour of the frameworks when all input are processed independently; i.e. each
task in the graph only depend on the previous one (see \ref{Figure fig:tg-inc}).

@@

% Histogram
\subsection{Histogram}
For our second experiment we find the frequency of the voxels intensity in BigBrain
and save them as a tuple in a file. In this experiment the end results depends on all
of the input results (see \ref{Figure fig:tg-histo}). This allows us to study the behaviour
of the framework when task are dependent. Also this application require data
shuffling thus inter worker communication. In this application we decided to omit the
dask futures api as it is less realistic for this use case.

@@

% BIDS Example
\subsection{BIDS Example}
Our last experiment compare both frameworks on the BIDS Example algorithm.
% Specify more details and the data set used
% SECTION TO COMPLETE -- Experience not done yet

@@

%% Incrementation
For the incrementation experiment, we
%% Node
first vary the number of instances in our cluster environment. We test with 1, 2, 4,
and 8 instances. From this we can evaluate the level of parallelism offered by
both framework.

@@

%% Splits
Finally, we seperate the BigBrain in different number of splits. We seperate it in
30, 125 and 750 splits; around 2.5GB, 0.6GB and 0.1GB, respectively. From that we can
evaluate the framework with task of different size and IO requirements.

%% Baseline
Our baseline for all experiment is 10 iterations, 4 sleep delay, 125 splits running
on 8 instances.

@@

%%% INCREMENTATION %%%
%% Instances
\subsection{Experiment 1: Number of instances}
Figure \ref{fig:inc_node}(a) shows that the incease in performance is not
proportional to the number of worker which differ form what we would expect; it even
decrease when we reach 8 workers. This is explained by an increase in IO overhead
corrolated to the number of workers (Figure \ref{fig:inc_node}(b)), as more worker
access the nfs simultaneously it increase the stress put on the system which result
in bottleneck, errors and slow down.

On the one hand, Dask seems better when using a single worker. On the other hand,
Spark makespan decrease faster for higher level of parallelism. However Dask is
constently faster then spark which make us think it has a better scheduler.

@@

%% Iterations
\subsection{Experiment 1: Number of iterations}
In Figure \ref{fig:inc_itr} we see that spark and the dask bag api are equivalent in
term of performance. The dask delayed api seems to perform well with low amount of
task but grow faster the spark and dask bag. We think this is because dask delayed
often schedule a block on different thread of the same worker which cause small delay
everytime due to inter thread data communication. This is good when there is fewer
tasks as is make the IO between block go out of sync hence lowering the IO bottleneck
however as the number of tasks increases those small delays become significant. The
dask futures api seems to outperform all the other apis but this is most likely
because the tasks are not inter dependent thus this application benifits from the
less optimal but faster scheduling brought by dask futures.

@@

%% Sleep time
\subsection{Experiment 1: Sleep delay}
As Figure \ref{fig:inc_sleep} 
% Not sure what to conclude.

@@

% Chunks
\subsection{Experiment 1: Number of splits}
In Figure \ref{fig:inc_splits} Spark is not compared for 30 chunks. This is because
spark has a 2GB limitation in the task size it can compute. Note from Figure
\ref{fig:inc_splits}(b) that the compute time increase when there is more splits.
This is expected because the sleep delay is constant throughout this experiment and
there are more compute task when the number of splits increases. We also note that in
Figure \ref{fig:inc_splits}(b) for 30 splits the dask bag api has a much lower
overhead time. This is because we only calculate the idle time of used core and the
dask bag api was only using one thread per block in comparison to the dask delayed
and dask futures api which were offloading the calculations on multiple thread of a
same worker. From Figure \ref{fig:inc_splits} we observe that when the dataset is
seperated in more (smaller) splits it saves IO time however it increases the overhead
by a similar amount; the inverse happens when we reduce teh number of splits.
